### 2021-04-15 supplementary.tex
label transfer

\documentclass[11pt, titlepage, twoside]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{authblk}

\makeatletter
\def\blx@maxline{77}
\makeatother

\captionsetup[figure]{labelfont=sc}
\captionsetup[table]{labelfont=sc}
\captionsetup[tabu]{labelfont=sc}

\DeclareCaptionType{MPEquation}[][List of equations]
\captionsetup[MPEquation]{labelformat=empty}

\DeclareCaptionType{MPListing}[][List of listings]
\captionsetup[MPListing]{labelformat=empty}

\usepackage{fancyhdr}
\setlength{\headheight}{16pt}
\pagestyle{fancyplain}
\fancyhf{}
%\fancypagestyle{plain}{}
\cfoot[\thepage]{\thepage}

\usepackage{stringenc}
\usepackage{pdfescape}

% Helper for debugging pdflatex errors induced by invalid UTF8 characters:
% output the actual character code causing the issue.

\makeatletter
\renewcommand*{\UTFviii@defined}[1]{%
\ifx#1\relax
\begingroup
% Remove prefix "\u8:"
\def\x##1:{}%
% Extract Unicode char from command name
% (utf8.def does not support surrogates)
\edef\x{\expandafter\x\string#1}%
\StringEncodingConvert\x\x{utf8}{utf16be}% convert to UTF-16BE
% Hexadecimal representation
\EdefEscapeHex\x\x
% Enhanced error message
\PackageError{inputenc}{Unicode\space char\space \string#1\space
(U+\x)\MessageBreak
not\space set\space up\space
for\space use\space with\space LaTeX}\@eha
\endgroup
\else\expandafter
#1%
\fi
}
\makeatother

% Make multi-paragraph footnotes align nicely on the left
\setlength\footnotemargin{10pt}

% Do not display an automatic Notes heading for endnotes, as they are contained
% by a Manuscripts section with a heading that has possibly been edited by the user
\def\enoteheading{}

% Apply footnote and endnote numbering schemes (decimal, lowercase roman etc.)
\renewcommand{\thefootnote}{\arabic{footnote}}
\renewcommand{\theendnote}{\arabic{endnote}}
% Adjust endnotes to use normal-sized numbering and text
\renewcommand{\enotesize}{\normalsize}
\renewcommand\enoteformat{%
  \raggedright
  \leftskip=1.8em
  \makebox[0pt][r]{\theenmark. \rule{0pt}{\dimexpr\ht\strutbox+\baselineskip}}%
}
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{document}
\title{Nested Stochastic Block Models Applied to the Analysis of Single Cell Data}
\author[1,2]{Leonardo Morelli}
\author[1,3]{Valentina Giansanti}
\author[1]{Davide Cittaro}
\affil[1]{Center for Omics Sciences, IRCCS San Raffaele Institute, Milan, Italy}
\affil[2]{Universit\`{a} Vita-Salute San Raffaele, Milan, Italy} 
\affil[3]{Department of Informatics, Systems and Communication, University of Milano-Bicocca, Milan, Italy}
\maketitle
\section*{Supplementary figures}

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_S1.png}
%\caption[]{Degree distribution of multiple \emph{k}NN graphs derived from scRNA-seq mixology datasets using variable number of Principal Components or number of neighbors. Each histogram shows the number of nodes (on y axis) within a specific degree bin (on x axis). Both the parameters influence the sparseness of the graph.}\label{MPFigure:F56BBDC3-0F0A-4558-8BC2-839E91250DB4}
%\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_S2.png}
%\caption[]{Adjusted Rand Index for different \emph{k}NN graphs after MCMC run. Maximal ARI over all hierarchy level is shown. Darker color indicates higher concordance with the ground truth}\label{MPFigure:9060BBC5-898E-4AA7-A57D-D221167F7250}
%\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_S3.png}
%\caption[]{Analysis of hematopoietic differentiation. (A) Low dimensional embedding of single cells colored by original cell type and pseudotime. (B) Cells are colored according to the nSBM grouping at level 3 of the hierarchy, next to a radial tree representation of the same model.}\label{MPFigure:6C191331-D4E9-4774-A8E0-210B2852D498}
%\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_Paul15_Leiden_r.png}
\caption[]{Low dimension embedding of single cells for hematopoietic differentiation colored according to Leiden clustering at decreasing resolution, from 1.0 to 0.1. Lowering the distribution does not grant that cells are grouped in a hierarchical way, \emph{e.g.\ } in the Erythroid branch groups 7 and 15 at resolution \emph{r}=1 are merged or split at coarser resolutions ($\gamma=0.6$ and $\gamma=0.3$)}\label{Figure_Paul15_Leiden_r}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{hist_label_second.png}
\caption[]{TO BE WRITTEN}\label{hist_label_second}
\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_S5.png}
%\caption[]{Adjusted Rand Index between cell clusters and cell type annotation filtering data at different cutoffs of consistency. Dot size is proportional to the number of cells remaining after filtering.}\label{MPFigure:5B291ABA-427A-42BE-BC26-38E626B1B088}
%\end{figure}

\end{document}