### 2021-07-18 supplementary.tex
added last figure

\documentclass[11pt, titlepage, twoside]{article}
\usepackage[top=20mm, bottom=20mm, left=20mm, right=20mm]{geometry}
\geometry{a4paper}
\geometry{portrait}
\usepackage[utf8]{inputenc}	% Comment out for xelatex; must use with pdflatex
\usepackage[T1]{fontenc}
\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{fontspec}
\usepackage{epstopdf}
\usepackage{float}
\usepackage[hang]{footmisc}
\usepackage{endnotes}
\usepackage{listingsutf8}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{tabu}
\usepackage{caption}
\usepackage{textcomp}
\usepackage{textgreek}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{authblk}
\usepackage{xr}
\externaldocument{manuscript}

\makeatletter
\def\blx@maxline{77}
\makeatother

\captionsetup[figure]{labelfont=sc}
\captionsetup[table]{labelfont=sc}
\captionsetup[tabu]{labelfont=sc}

\DeclareCaptionType{MPEquation}[][List of equations]
\captionsetup[MPEquation]{labelformat=empty}

\DeclareCaptionType{MPListing}[][List of listings]
\captionsetup[MPListing]{labelformat=empty}

\usepackage{fancyhdr}
\setlength{\headheight}{16pt}
\pagestyle{fancyplain}
\fancyhf{}
%\fancypagestyle{plain}{}
\cfoot[\thepage]{\thepage}

\usepackage{stringenc}
\usepackage{pdfescape}

% Helper for debugging pdflatex errors induced by invalid UTF8 characters:
% output the actual character code causing the issue.

\makeatletter
\renewcommand*{\UTFviii@defined}[1]{%
\ifx#1\relax
\begingroup
% Remove prefix "\u8:"
\def\x##1:{}%
% Extract Unicode char from command name
% (utf8.def does not support surrogates)
\edef\x{\expandafter\x\string#1}%
\StringEncodingConvert\x\x{utf8}{utf16be}% convert to UTF-16BE
% Hexadecimal representation
\EdefEscapeHex\x\x
% Enhanced error message
\PackageError{inputenc}{Unicode\space char\space \string#1\space
(U+\x)\MessageBreak
not\space set\space up\space
for\space use\space with\space LaTeX}\@eha
\endgroup
\else\expandafter
#1%
\fi
}
\makeatother

% Make multi-paragraph footnotes align nicely on the left
\setlength\footnotemargin{10pt}

% Do not display an automatic Notes heading for endnotes, as they are contained
% by a Manuscripts section with a heading that has possibly been edited by the user
\def\enoteheading{}

% Apply footnote and endnote numbering schemes (decimal, lowercase roman etc.)
\renewcommand{\thefootnote}{\arabic{footnote}}
\renewcommand{\theendnote}{\arabic{endnote}}
% Adjust endnotes to use normal-sized numbering and text
\renewcommand{\enotesize}{\normalsize}
\renewcommand\enoteformat{%
  \raggedright
  \leftskip=1.8em
  \makebox[0pt][r]{\theenmark. \rule{0pt}{\dimexpr\ht\strutbox+\baselineskip}}%
}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}
\title{Nested Stochastic Block Models Applied to the Analysis of Single Cell Data}
\author[1,2]{Leonardo Morelli}
\author[1,3]{Valentina Giansanti}
\author[1]{Davide Cittaro}
\affil[1]{Center for Omics Sciences, IRCCS San Raffaele Institute, Milan, Italy}
\affil[2]{Universit\`{a} Vita-Salute San Raffaele, Milan, Italy} 
\affil[3]{Department of Informatics, Systems and Communication, University of Milano-Bicocca, Milan, Italy}
\maketitle

\section*{Supplementary tables}
\begin{table}[h!]
\centering
 \begin{tabular}{||c|c c|c c||}
%  \begin{tabular}{|| m{7em} | m{5em} | m{7em} | m{7em} | m{5em} ||}
 \hline
\multicolumn{1}{||c|}{\textbf{Noise level ($\sigma$)}} & \multicolumn{2}{c|}{\textbf{Leiden}} & \multicolumn{2}{c||}{\textbf{nSBM level 1}} \\

 & $ARI$ & $N$ & $ARI$ & $N$  \\ 
 \hline\hline
-- & 0.9228 & 8 & 0.8944 & 6 \\
\hline 
0.5 & 0.6249 & 10 & 0.7509  & 10 \\ 
 \hline
 1.0  & 0.5966 & 7 & 0.7269  & 6 \\
 \hline
 1.5  & 0.2833 & 9 &0.3029 & 3 \\
 \hline
 2.0  & 0.1405 & 12 & 0.1961 & 2 \\
 \hline
\end{tabular}
\caption{\textcolor{red}{Effect of random noise }}
\label{Table_Noise}
\end{table}

\clearpage

\section*{Supplementary figures}

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.375\textheight]{Random_Leiden_by_gamma.png}
\caption[]{\textcolor{red}{foobarbaz}}
\label{Modularity_Random_Res}
\end{figure}
\clearpage

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Tian_NSBM0.png}
%\caption[]{Analysis of a mixture of 5 known cell lines. (A) UMAP  embedding of cell lines from the sc-mixology experiment coloured by the level 0 of the hierarchy proposed by the nested Stochastic Block Model (B) UMAP embedding coloured by the classification made by SCCAF when partitions in (A) are used.}\label{Tian_NSBM0}
%\end{figure}
%\clearpage
%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_S2.png}
%\caption[]{Adjusted Rand Index for different \emph{k}NN graphs after MCMC run. Maximal ARI over all hierarchy level is shown. Darker color indicates higher concordance with the ground truth}\label{MPFigure:9060BBC5-898E-4AA7-A57D-D221167F7250}
%\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_Hemato_Supp.pdf}
\caption[]{Analysis of hematopoietic differentiation. (A) Low dimensional embedding of single cells coloured by original cell type and pseudotime. (B) Cells are coloured according to the nSBM grouping at level \textcolor{red}{1} of the hierarchy, next to a radial tree representation of the same model.}\label{Figure_Hemato_Supp}
\end{figure}
\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{Figure_Paul15_Leiden_r.pdf}
\caption[]{Low dimension embedding of single cells for hematopoietic differentiation coloured according to Leiden clustering at decreasing resolution, from 1.0 to 0.1. Lowering the distribution does not grant that cells are grouped in a hierarchical way, \emph{e.g.} groups 6 and 8 in the Erythroid branch at resolution $\gamma=1$ are merged or split at coarser resolutions ($\gamma=0.6$ and $\gamma=0.3$).}\label{Figure_Paul15_Leiden_r}
\end{figure}
\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,width=0.50\textheight]{Figure_Stability.png}
\caption[]{Cell Stability. UMAP embeddings of PBMC data profiled with four different technologies ranked by their quality. Cells are coloured by the Cell Stability metric.}\label{Figure_Stability}
\end{figure}
\clearpage

%\begin{figure}[htbp]
%\centering
%\includegraphics[keepaspectratio,width=\textwidth,width=0.50\textheight]{Figure_DE.png}
%\caption[]{Results of differential gene expression with or without marginals. The heatmaps show the difference in sensitivity (left) or recall (right) of statistical testing for differential genes among cell groups when marginals are included or not in the model. Tests have been conducted on simulated data with different levels of noise (BCV) or probability of differential expression (p(DE))}\label{Figure_DE}
%\end{figure}
%\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=\textwidth,height=0.75\textheight]{hist_label_second.png}
\caption[]{Distribution of second choices in label transfer. For each dataset in which we assessed accuracy of label transfer by SBM, we collect the rank of affinities of the correct class when an assignment could not be performed (\emph{i.e.} the highest affinity was for the "Unknonwn" label)  }\label{hist_label_second}
\end{figure}
\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[keepaspectratio,width=1\textwidth,height=\textheight]{FigureOptimalLevel.pdf}
\vspace*{-80pt}
\caption[]{Choice of the optimal hierarchy level. Datasets analysed for choice of the most relevant hierarchy level are represented here, from top to bottom in the same order as in Table \ref{table_optimal}. For each dataset we report the UMAP embedding coloured by the annotation given in the corresponding manuscript, the profile of modularity $Q$ at different level of the hierarchy and the UMAP embeddings coloured by the level according to $i^k$ or $i^Q$. The dashed line in modularity plots represents the modularity calculated using the annotation from the corresponding manuscripts.}\label{FigureOptimalLevel}
\end{figure}

\clearpage
\begin{figure}[H]
\centering
\includegraphics[keepaspectratio,width=1\textwidth,height=\textheight]{Figure_Rare_Cells_Supp.pdf}
%\vspace*{-40pt}
\caption[]{Identification of rare populations. FOOBAR} \label{Figure_Rare_Supp}
\end{figure}

\clearpage
\begin{figure}[H]
\centering
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel0.pdf} 
        \caption{Chromium 10x} \label{cm_10x}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel1.pdf} 
        \caption{iCell8} \label{cm_icell8}
    \end{subfigure}
    \hfill
%    \vspace{40pt}
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel2.pdf} 
        \caption{Quartz-seq2} \label{cm_quartz}
    \end{subfigure}

    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel3.pdf} 
        \caption{MARS-seq} \label{cm_mars}
    \end{subfigure}
%    \vspace{40pt}
   \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel4.pdf} 
        \caption{Crypt cells} \label{cm_crypt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ClusterMap_L0_Panel5.pdf} 
        \caption{Planaria} \label{cm_planaria}
    \end{subfigure}
    \vspace{40pt}

%\vspace*{-40pt}
\caption[]{Consistency at Level 0} \label{Figure_Clustermap_L0}
\end{figure}

\end{document}